Manuscript title here

### 2024-09-11 manuscript.Rmd
sparar för natten
Lagt till retts och news i stycke 3, men kasnke ska flytta delar till metod?

Test

@@

@@

Shock are generally defined as circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are five types of causes, which are hypovolemia, obstructive, neurogenic, cardiogenic and distributive. These causes can occur alone or in combination. The most usuall one during trauma are hypovolemia in form of a large bleeding[@teixeira_preventable_2007].

Clinically to be able to differentiate the degree of shock, the ATLS classification for major haemorrhage can be used. The classification are divided into four classes, based on the patients: estimated bloodloss, heart rate, bloodpressure, pulse pressure, respiratory rate, mental status and urine output. However this only suits hypovolemic shocks and not the other types.

Skriva om stycke 4 och istället lägga in innehåll om använda retts som en fortsättnig på stycke 3.

  (4  Kliniskt för att förbättra omhändertagande används OFI
    Förklara vad OFI är för något

  To be able to asses and improve the quality of trauma care, improvement programmes aim to identify opportunities for improvements(OFI). OFI is multidiciplinary morbidity and mortality reviews, where oppertunity for improvemetns are identified in the structure and clinical process of trauma care(Hussein).)

@@

(a) Logistic regression to determine the association between the binary results of OFI existence and non-binary degree of shock.

@@

>>>>

@@

  (4  Kliniskt för att förbättra omhändertagande används OFI
    Förklara vad OFI är för något
      To be able to asses and improve the quality of trauma care, improvement programmes aim to identify opportunities for improvements(OFI). OFI is multidiciplinary morbidity and mortality reviews, where oppertunity for improvemetns are identified in the structure and clinical process of trauma care(Hussein).)

@@

Shock are generally defined as circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are four types of causes, which are hypovolemic, obstructive, cardiogenic and distributive. These causes can occur alone or in combination. The most usual one during trauma are hypovolemic in form of a large bleeding[@teixeira_preventable_2007].

Clinically to be able to differentiate the degree of shock, the ATLS classification for major haemorrhage can be used. The classification are divided into four classes, based on the patients: estimated bloodloss, heart frequency (HF), bloodpressure (BP), pulse pressure, respiratory rate (RR), mental status and urine output. However this only suits hypovolemic shocks and not the other types. There are two other classifications, namely RETTS and NEWS2. These are used in emergency triaging and can be used on other types of shocks, however they are not shock specific classifications. If we look into all three classifications we will be able to find similarities in that all use HF, BP, RR and mental status. Therefore we decided to use these markers as the independent variables for classification of shock regardless of their nature. (sista två meningarna borde tillhöra metod...)

@@

(a) Logistic regression to determine the association between the binary results of OFI existence and teh four degrees of shock.

@@

```{r main, echo=FALSE} 
source("main.R") ## This "imports" the main script file of your project and run any code in it 
```
